Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-09-23 paper/main.tex
add statement about axiliary nature of rigidity and spring tension

In addition to springs between individual vertices $\vec{x}_i$, we assume that each angle at a
vertex between edges is subject to a force indiced by curvature.
% We assume that we can model the mechanical properties of the bacterium as an elastic rod.
Given the angle $\alpha_i = \sphericalangle(\vec{c}_{i-1},\vec{c}_i)$ between adjacent edges,
the bending force is proportional to the curvature $\kappa_i$ at each vertex
$\kappa_i = 2\tan(\alpha_i/2)$.
The resulting force acts along the angle bisector which can be calculated from the edge vectors.
Figure~\ref{fig:mechanics-bacterium} shows an elongated rod represented by $6$ vertices.
The forces acting on vertices $\vec{x}_i,\vec{x}_{i-1},\vec{x}_{i+1}$ are given by

@@

We chose to optimize for the damping of agents' movement within the medium, 3 interaction parameters
of the Mie potential $n,m,V_0$ a single radius for all agents and growth rates which were chosen
individually for each cell.
We removed any stochastic motion contributions by setting $D=0$.
This leads to a total of $5+1\times 6=11$ variable parameters.
In order to model non-flexible rods, we set the rigidity and spring tension of the rod to a high,
fixed value.
To estimate these parameters we used the differential evolution~\cite{Storn1997} global optimization
algorithm which was initially seeded with a sample generated by the latin-hypercube~\cite{McKay1979}
method.
Finally, we performed a local minimization using the L-BFGS-B algorithm~\cite{Liu1989}.

>>>>

The value of the spring constants determines the elasticity with respect to elongation of the cell.
In situations where this aspect is not relevant, they merely serve as a auxiliary parameters to
maintain the length of the rod.
In addition to springs between individual vertices $\vec{x}_i$, we assume that each angle at a
vertex between edges is subject to a force indiced by curvature.
% We assume that we can model the mechanical properties of the bacterium as an elastic rod.
Given the angle $\alpha_i = \sphericalangle(\vec{c}_{i-1},\vec{c}_i)$ between adjacent edges,
the bending force is proportional to the curvature $\kappa_i$ at each vertex
$\kappa_i = 2\tan(\alpha_i/2)$.
The resulting force acts along the angle bisector which can be calculated from the edge vectors.
Figure~\ref{fig:mechanics-bacterium} shows an elongated rod represented by $6$ vertices.
The forces acting on vertices $\vec{x}_i,\vec{x}_{i-1},\vec{x}_{i+1}$ are given by

@@

We chose to optimize for the damping of agents' movement within the medium, 3 interaction parameters
of the Mie potential $n,m,V_0$ a single radius for all agents and growth rates which were chosen
individually for each cell.
We removed any stochastic motion contributions by setting $D=0$.
This leads to a total of $5+1\times 6=11$ variable parameters.
As highlighted before, in order to model non-flexible rods, we set the rigidity and spring tension
of the rod to a high, fixed value.
To estimate these parameters we used the differential evolution~\cite{Storn1997} global optimization
algorithm which was initially seeded with a sample generated by the latin-hypercube~\cite{McKay1979}
method.
Finally, we performed a local minimization using the L-BFGS-B algorithm~\cite{Liu1989}.
